
# Satellite cubes

Digital Earth Australia [DEA](https://www.dea.ga.gov.au/) provides earth observation data for Australia. DEA data was accessed via a workflow that:

* finds features available for an area and timeframe
* creates an image collection from available features
* defines a [regular cube](https://e-sensing.github.io/sitsbook/earth-observation-data-cubes.html#regular-image-data-cubes) structure
* create a cloud mask
* merge the image collection into the parameters defined by the regular cube, taking into account the cloud mask

The `envRas` project makes use of the `envRaster` package `r cite_package("envRaster")` to abstract all these steps into the single function `make_sat_cube`. 

## Example: building a single seasonal image

The following example finds and processes Landsat 8 data to generate median Normalized Difference Vegetation Index ([NDVI](https://www.usgs.gov/landsat-missions/landsat-normalized-difference-vegetation-index)) for the `r settings$name` example area in summer 2019/2020. The same process is repeated to generate seasonal images to the temporal extent desired (or allowed by the satellite collections).

Features available within DEA data were found using [DEA’s SpatioTemporal Asset Catalogue (STAC) metadata](https://knowledge.dea.ga.gov.au/guides/setup/gis/stac/). [STAC](https://stacspec.org/en) is a specification that provides a common language to describe geospatial information so it can more easily be discovered. The `rstac` package `r cite_package("rstac")` was used to find the images available for each DEA collection specified for an area of interest.

A bounding box is required to define the area for `rstac` to search. In this case, the coordinate systems for the bounding box should be in WGS84 decimal degrees ([epsg: `4326`](https://epsg.io/4326)), irrespective of the desired spatial coordinate system of the end result.

The next step is to create an image collection using the function `stac_image_collection` from the `gdalcubes` package `r cite_package("gdalcubes")`. In the example below, only images with less than 10% cloud cover are returned, supplied in the example below as `property_filter`. Other arguments can be passed via the `dots` argument in `make_sat_cube` to `stac_image_collection`.

A regular cube is then defined via the `gdalcubes` function `cube_view`. Here an extent is required: a bounding box with the addition of bounding dates. In this case the coordinate system for the extent should be in the desired coordinate system for the resulting cube ([epsg: `r settings$epsg_proj`](`r paste0("https://epsg.io/", settings$epsg_proj)`)). The `make_sat_cube` takes the spatial details from the base_grid supplied and the temporal details from the period supplied.

A cloud mask is created via the `mask` argument to `make_sat_cube`, in this case specifying that pixels identified as cloud and cloud shadow should be masked from the results.

Finally, the collection, regular cube and cloud mask are brought together, saving the file locally.

```{r satExample, include = TRUE, echo = TRUE}

  start_date <- "2019-12-01"
  end_date <- "2020-02-29"
  
  base_grid <- terra::rast(extent = settings$boundary
                           , crs = terra::crs(settings$boundary)
                           , res = c(settings$sat_res, settings$sat_res)
                           , vals = 1
                           )
  
  out_file <- fs::path(settings$out_report, "out", "ndvi__summer__2019-12-01.tif")
  
  if(!file.exists(out_file)) {

    make_sat_cube(base_grid
                  , start_date = start_date
                  , end_date = end_date
                  , season = "summer"
                  , out_dir = dirname(out_file)
                  , collections = c("ga_ls8c_ard_3", "ga_ls9c_ard_3")
                  , period = "P3M"
                  , layers = NULL
                  , indices = list("ndvi" = c("nir", "red"))
                  , mask = list(band = "oa_fmask", mask = c(2, 3))
                  , chunks = c(1, 1)
                  , chunk_dir = fs::path(Sys.getenv("TMPDIR"), "sat_cube_temp")
                  , chunk_prefix = "tile_"
                  , sleep = 60
                  , attempts = 1
                  , property_filter = \(x) {x[["eo:cloud_cover"]] < 10}
                  )
    
  }

```

Figure \@ref(fig:examplePlot) shows the Normalized Difference Vegetation Index ([NDVI](https://www.usgs.gov/landsat-missions/landsat-normalized-difference-vegetation-index))) result for the example area and timeframe.

```{r examplePlot, fig.cap = "NDVI result for the example area"}
  
  terra::rast(out_file) %>%
    terra::mask(mask = terra::vect(settings$boundary)) %>%
    terra::plot()

```

## Example: building an epoch cube from seasonal cube

Once each season and band of a seasonal cube is available, an 'epoch' cube can be built from each season within the epoch. In the example below, the 10 year median values are generated from the previously built seasonal cube.

```{r satExampleEpoch, include = TRUE, echo = TRUE}

  # Information needed to build an epoch cube
  epochs <- envFunc::make_epochs(start_year = settings[["start_year", exact = TRUE]]
                                 , end_year = settings[["end_year", exact = TRUE]]
                                 , epoch_step = settings[["epoch_period", exact = TRUE]]
                                 , epoch_overlap = FALSE
                                 ) %>%
    tidyr::unnest(cols = c(years)) %>%
    dplyr::rename(year = years) %>%
    dplyr::select(year, epoch) %>%
    dplyr::left_join(settings[["seasons", exact = TRUE]]$seasons) %>%
    dplyr::mutate(start_date = as.character(start_date)) %>%
    dplyr::inner_join(name_env_tif(settings$sat_seas_cube_dir, parse = TRUE) %>%
                        dplyr::filter(layer == "ndvi")
                      ) %>%
    dplyr::filter(!is.na(path)) %>%
    dplyr::mutate(scale = dplyr::case_when(layer %in% names(settings$sat_indices) ~ gdalcubes::pack_minmax(min = -1, max = 1)$scale
                                            , !layer %in% names(settings$sat_indices) ~ gdalcubes::pack_minmax(min = 0, max = 10000)$scale
                                            , TRUE ~ 1
                                            )
                  , offset = dplyr::case_when(layer %in% names(settings$sat_indices) ~ gdalcubes::pack_minmax(min = -1, max = 1)$offset
                                            , !layer %in% names(settings$sat_indices) ~ gdalcubes::pack_minmax(min = 0, max = 10000)$offset
                                            , TRUE ~ 0
                                            )
                  ) %>%
    tidyr::nest(data = c(year, start_date, end_date, path)) %>%
    dplyr::mutate(start_date = purrr::map_chr(data, \(x) as.character(min(x$start_date)))) %>%
    name_env_tif() %>%
    dplyr::mutate(out_file = fs::path("I:"
                                      , gsub("P3M", "P10Y", out_file)
                                      )
                  ) %>%
    dplyr::mutate(stack = purrr::map(data
                                     , ~ terra::rast(.$path)
                                     )
                  , done = file.exists(out_file)
                  )

  # Build the epoch cube
  fs::dir_create(dirname(epochs$out_file))
  
  purrr::pwalk(list(epochs$stack[!epochs$done]
                    , epochs$out_file[!epochs$done]
                    , epochs$scale[!epochs$done]
                    , epochs$offset[!epochs$done]
                    )
               , \(a, b, c, d) terra::app(a
                                          , fun = "median"
                                          , na.rm = TRUE
                                          , filename = b
                                          , overwrite = TRUE
                                          , wopt = list(datatype = "INT2S"
                                                        , scale = c
                                                        , offset = d
                                                        , gdal = c("COMPRESS=NONE")
                                                        )
                                          )
               )


```

Figure \@ref(fig:examplePlotEpoch) shows the Normalized Difference Vegetation Index ([NDVI](https://www.usgs.gov/landsat-missions/landsat-normalized-difference-vegetation-index))) result for the example area and timeframe.

```{r examplePlotEpoch, fig.cap = "NDVI result for the example area over a 10 year period"}
  
  ex_data <- name_env_tif(epochs %>% dplyr::select(path = out_file), parse = TRUE) %>%
    dplyr::mutate(name = paste0(layer, " ", season))

  ex <- terra::mask(terra::rast(ex_data$path), mask = terra::vect(settings$boundary))

  names(ex) <- ex_data$name
  
  terra::plot(ex
              , nc = 1
              )

```

## Satellite seasonal cube

```{r satSeasCube}

  sat_seas_cube_data <- name_env_tif(settings$sat_seas_cube_dir
                                , parse = TRUE
                                ) %>%
    dplyr::filter(source == "DEA")

  sat_layers <- sat_seas_cube_data  %>%
    tidyr::separate_longer_delim(collection, "--") %>%
    dplyr::left_join(ras_collection) %>%
    dplyr::mutate(start_date = as.Date(start_date)
                  , collection_md = paste0("[", collection, "](", coll_url, ")")
                  ) %>%
    dplyr::group_by(source, season) %>%
    dplyr::summarise(collection = stringr::str_flatten_comma(unique(collection_md))
                     , starts = stringr::str_flatten_comma(unique(start_date))
                     , layers = stringr::str_flatten_comma(unique(layer))
                     ) %>%
    dplyr::ungroup() %>%
    dplyr::select(source, collection, season, starts, layers)

```

Table \@ref(tab:satCollectionTab) shows the collection, layers and time periods prepared from DEA data for the seasonal cube. This created a regular cube of `r nrow(sat_seas_cube_data)` files made up of `r length(unique(sat_seas_cube_data$layer))` layers with seasonal data for each of `r length(unique(sat_seas_cube_data$season))` seasons in `r length(unique(year(sat_seas_cube_data$start_date)))` years. This seasonal cube had the following parameters:

* first time period: `r min(sat_seas_cube_data$start_date)`
* last time period: `r max(sat_seas_cube_data$start_date)`
* period (in [ISO-8601 duration](https://www.digi.com/resources/documentation/digidocs/90001488-13/reference/r_iso_8601_duration_format.htm) format): `r settings$period`
* x and y distance: `r settings$sat_res` m
* seasons: `r sat_seas_cube_data$season %>% unique %>% vec_to_sentence`
* December data summarized into following year. e.g. December 2019 data will appear in the summer 2020 layer

[Appendix][appendices] Table \@ref(tab:satSeasCubeApp) lists details for each of the existing layers in the seasonal cube.

<br>

```{r satCollectionTab}  
  
  knitr::kable(sat_layers
               , caption = "DEA collections and layers prepared by envRas"
               )

```

## Satellite epoch cube

```{r satEpochCube}

  sat_epoch_cube_data <- epoch_cube_data %>%
    dplyr::filter(grepl("DEA", source))

```

A cube with `r settings$epoch_period`-year epochs, was then generated from the [seasonal cube][satellite seasonal cube]. The epoch cube had the same parameters as the seasonal cube, except that the period was `P10Y` instead of `P3M`. Note that, say, summer, in the `r settings$epoch_period`-year epoch cube took the median values for each summer image in the seasonal cube.

Table \@ref(tab:satEpochTab) shows a summary of the layers and layer values.

```{r satEpochTab}

  knitr::kable(sat_epoch_cube_data %>%
          dplyr::mutate(across(where(is.numeric), ~round(.x, 2))
                        , iqr = paste0(q25
                                     , " to "
                                     , q75
                                     )
                        ) %>%
          dplyr::select(layer
                        , season
                        , median = q50
                        #, sd
                        , `inter-quartile range` = iqr
                        ) %>%
          janitor::remove_empty(which = "cols")
        , caption = paste0("Summary of "
                           , unique(sat_epoch_cube_data$source)
                           , " (collection "
                           , unique(sat_epoch_cube_data$collection)
                           , ") layers for the "
                           , settings$name
                           , " example area"
                           )
        , align = "llrr"
        )

```

```{r satEpochRmd, eval = do_sat_summary, include = FALSE}

  sat_layers <- unique(sat_epoch_cube_data$layerid)

  sat_layer_rmd = NULL
  
  for (i in sat_layers) {
    
    this_layer <- i
    
    sat_layer_rmd = c(sat_layer_rmd
                      , knitr::knit_expand(here::here("report", "child", "layer_rmd.Rmd"))
                      )
    
  }
  
```

`r if(do_sat_summary) paste(knitr::knit(text = sat_layer_rmd), collapse = '\n')`
