
# Climate cubes

ANUClimate 2.0 [@RN4594] is available from the National Computing Infrastructure ([NCI](https://nci.org.au/)) via their [THREDDS](https://docs.unidata.ucar.edu/tds/4.6/adminguide/) service [@RN4593].

NCI data was accessed via a workflow that:

* lists files for each season using the naming structure in the THREDDS service
    + each season had three files, one for each month in the season
* created a connection to each of those files
* subset the data to the desired geographic extent
* reduced the time axis to a single pixel for that season (min, mean, or max, depending on climate variable)
* adjusted the result to a grid in the desired output coordinate reference system
* saved the result locally

## Example

There is no STAC specification for the ANUClimate data available from the NCI. Instead, a series of filenames are built up using a pattern. Each file is the monthly (or daily are also available) climate data for all of Australia at roughly 1 km grid size. Each file is then queried for the geographic subset required.

As per the `get_sat_data`, there is an equivalent `get_cli_data` function in the `envRaster` package that abstracts many of the steps to download a climate cube.

```{r cliExampleFiles}

  cli_example_files <- cli_files %>%
    dplyr::filter(start_date == as.Date(start_date)) %>%
    dplyr::mutate(out_file = here::here("report", "out", "cli_example.tif"))
  
  cli_example_files$data[[1]]
  
```

Three files representing monthly rainfall for the summer of 2014 (December 2013 to February 2014).

```{r cliExampleGet, cache = TRUE}
  
  purrr::pwalk(list(cli_example_files$data[[1]]
                    , cli_example_files$out_file[[1]]
                    , cli_example_files$func[[1]]
                    , cli_example_files$scale[[1]]
                    , cli_example_files$offset[[1]]
                    )
               , get_cli_data
               , base = base_grid
               )
  
```

The `get_cli_data` function does the work of saving the resulting cube to the file specified in `cli_examples_files` object `r cli_example_files$out_file`. That example raster can be seen in Figure \@ref(fig:cliExampleMap).

```{r cliExampleMap, fig.cap = "Summer rain result from the call to `get_cli_data`"}

  plot(terra::rast(cli_example_files$out_file[[1]]))

```

## Climate seasonal cube

```{r cliSeasonalCubeData}

  cli_seas_cube_dir <- fs::path("H:"
                                , "data"
                                , 'raster'
                                , "cube__P3M"
                                , "NCI__ANUClimate__sa_ibrasub_xn____0"
                                )

  cli_seas_cube <- parse_env_tif(cli_seas_cube_dir
                           , parse = TRUE
                           ) %>%
    dplyr::filter(source == "NCI") %>%
    dplyr::mutate(start_date = as.Date(start_date))

  cli_layers <- cli_seas_cube %>%
    dplyr::select(source, collection, band) %>%
    dplyr::distinct()

  cli_seasons <- settings$epochs %>%
    tidyr::unnest(cols = c(years)) %>%
    dplyr::rename(year = years) %>%
    dplyr::left_join(settings$seasons$seasons) %>%
    dplyr::inner_join(cli_seas_cube) %>%
    dplyr::select(year, epoch, season, start_date, end_date) %>%
    dplyr::distinct()

```

Table \@ref(tab:cliCollectionTab) shows the collection and layers prepared from NCI data for each of the seasons in Table \@ref(tab:cliSeasonsTab). This created a regular cube of `r nrow(cli_seas_cube)` files made up of `r nrow(cli_layers)` layers with seasonal data for each of `r nrow(cli_seasons)` seasons. This seasonal cube had the following parameters:

* start date: `r min(cli_seasons$start_date)`
* end date: `r max(cli_seasons$end_date)`
* period (in [ISO-8601 duration](https://www.digi.com/resources/documentation/digidocs/90001488-13/reference/r_iso_8601_duration_format.htm) format): `r settings$period`
* x and y distance: `r settings$cli_res` m
* seasons: `r cli_seasons$season %>% unique %>% vec_to_sentence`
* December data summarized into following year. e.g. December 2019 data will appear in the summer 2020 layer

[Appendix][appendices] Table \@ref(tab:cliSeasCubeApp) lists details for each of the existing layers in the seasonal cube.

<br>

```{r cliCollectionTab}  
  
  knitr::kable(cli_layers %>%
                      dplyr::mutate(collection = gsub("--", " and ", collection))
                    , caption = "NCI collections and layers prepared by envRas"
                    )

```

<br>

```{r cliSeasonsTab}

  knitr::kable(cli_seasons
                    , caption = "Start and end dates defining seasons for preparing NCI layers"
                    )

```

## Climate epoch cube

```{r cliEpochCubeData}

  cli_epoch_cube_dir <- fs::path("H:", "data", "raster", "sa_ibrasub_xn______0")

  cli_tifs <- name_env_tif(cli_epoch_cube_dir, parse = T) %>%
    dplyr::filter(grepl("NCI", path))

  cli_epoch_cube <- terra::rast(cli_tifs$path)

  cli_epoch_cube_data <- cli_tifs %>%
    dplyr::mutate(values = purrr::map(path, \(x) terra::values(terra::rast(x)))
                  , data = purrr::map(values, \(x) quibble(x, na.rm = TRUE))
                  , use_name = paste0(layer, " ", season)
                  , layerid = gsub("[[:punct:]]", "", layer)
                  ) %>%
    tidyr::unnest(cols = c(data))
  
  names(cli_epoch_cube) <- cli_epoch_cube_data$use_name

```

A cube with `r settings$epoch_period`-year epochs, was then generated from the [seasonal cube]. The epoch cube had the same parameters as the seasonal cube, except that the period was `P10Y` instead of `P3M`. Note that, say, summer, in the `r settings$epoch_period`-year epoch cube took the median values for each summer image in the seasonl cube.

Table \@ref(tab:cliEpochTab) shows a summary of the layers and layer values.

```{r cliEpochTab}

  knitr::kable(cli_epoch_cube_data %>%
                 dplyr::mutate(across(where(is.numeric), ~round(.x, 2))
                               , iqr = paste0(q25
                                              , " to "
                                              , q75
                                              )
                               ) %>%
                 dplyr::select(layer
                               , season
                               , median = q50
                               #, sd
                               , `inter-quartile range` = iqr
                               ) %>%
                 janitor::remove_empty(which = "cols")
               , caption = paste0("Summary of "
                                  , unique(cli_epoch_cube_data$source)
                                  , " (collection "
                                  , unique(cli_epoch_cube_data$collection)
                                  , ") layers for the "
                                  , settings$name
                                  , " example area"
                                  )
               )

```

```{r cliEpochRmd, eval = do_cli_summary, include = FALSE}

  cli_layers <- unique(cli_epoch_cube_data$layerid)

  cli_layer_rmd = NULL
  
  for (i in cli_layers) {
    
    this_layer <- i
    
    cli_layer_rmd = c(cli_layer_rmd
                      , knitr::knit_expand(here::here("report", "child", "cli_layer_rmd.Rmd"))
                      )
    
  }
  
```

`r if(do_cli_summary) paste(knitr::knit(text = cli_layer_rmd), collapse = '\n')`
