```{r {{this_data_name}}}

  use_value <- tolower("{{this_data_name}}") # use in figure references and file names

```

```{r dataSetup{{this_data_name}}}
library(dplyr)
library(envReport)
library(lubridate)
library(ggplot2)
library(sf)

bio_all_file <- "H:/data/vector/bio_all.parquet"
use_value <- "alis"

settings <- rio::import("H:/projects/envPIA/out/analysis/pia__50000__spatial__90__Study_Area__Study Area A1--Study Area B2--Study Area D--Study Area C--Study Area A2--Study Area B1/settings.rds")

  codes <- fs::dir_ls(fs::path(here::here()
                               , "function"
                               )
                      ) %>%
    grep("\\/_", ., value = TRUE, invert = TRUE)
  
  # safe_source <- safely(source) # try not to use this as it is better to error here than half way through the workflow
  
  purrr::walk(codes
              , source
              )


bio_data_name_dir <- "H:/projects/envOcc/out/ds"

  #setup aligning col names
  lunames <- envImport::data_map %>% 
    dplyr::filter(data_name == use_value) %>% 
    dplyr::select(-epsg, -order, -days) %>% 
    tidyr::pivot_longer(names_to = "name", cols = -data_name, values_to = "fromdata") %>% 
    dplyr::select(name, fromdata) %>% 
    tidyr::drop_na()


  #collect source's data
  bio_file <- fs::path(bio_data_name_dir, use_value, paste0(use_value, "_raw.rds"))

  # if(file.exists(bio_file)){
  # 
  # bio_df <- rio::import(bio_file)
  # 
  # } else {
      
      bio_df <- arrow::open_dataset(bio_all_file) %>%
        dplyr::filter(year >= settings$min_year) %>%
        dplyr::filter(data_name == use_value) %>%
        dplyr::filter(original_name != "") %>% 
        dplyr::collect()
      
      # }
    
  # plots setup
      if(use_value != "alis"){
  bio_df <- bio_df %>% 
    dplyr::select(any_of(lunames$fromdata)) %>%
    plyr::rename(setNames(lunames$name, lunames$fromdata))
      }
  
  include_time_plot <- bio_df %>%
    dplyr::mutate(year = lubridate::year(date)
                  , month = lubridate::month(date)
    ) %>%
    dplyr::pull(year) %>%
    n_distinct() > 1
  
  
    include_summarise_cols <- all(c("original_name", "survey", "kingdom", "rel_metres", "month", "year") %in% names(bio_df))

    
  # extract_date <- data_info_bio %>%
  #   dplyr::filter(data_name == use_value) %>%
  #   {if(nrow(.) > 1) (.) %>% dplyr::mutate(text_date = paste0(text_date
  #                                                             , " ("
  #                                                             , if_else(kingdom == "Plantae", "flora", "fauna")
  #                                                             , ")"
  #                                                             )
  #                                          )  else (.)
  #     } %>%
  #   dplyr::pull(text_date) %>%
  #   envFunc::vec_to_sentence()
  

 
```

(ref:timePlotReference{{this_data_name}}) Figure \@ref(fig:timePlot{{this_data_name}})

#### `r paste0(use_value, " (", use_value, ") {#", toupper(use_value), "}")`

##### Background
```{r dataNameSummary{{this_data_name}}, child = here::here("report", "child", "sources", paste0(use_value, ".Rmd"))}
```

<br>

##### Data

`r if(include_summarise_cols) {summarise_cols(bio_df, c("original_name", "survey", "kingdom", "rel_metres", "month", "year"), qs = c(0.05, 0.25, 0.5, 0.75, 0.95))$text}`

`r if(include_time_plot) {paste0("(ref:timePlotReference{{this_data_name}}) shows the distribution of records through time for ",use_value,". ")}`Figure \@ref(fig:dataNameMap{{this_data_name}}) shows the spatial location of records for `r use_value`. The region boundaries used in Figure \@ref(fig:dataNameMap{{this_data_name}}) are those of the Regional Species Conservation Assessment ((RSCA)[#rsca]) project (also see in Figure \@ref(fig:rscalsaFig)).

<br>

```{r timePlot{{this_data_name}}, fig.cap = paste0("Records from ",use_value," through time"), eval = include_time_plot}

  data_name_time_plot(bio_df)

```

`r if(include_time_plot) {"<br>"}`

```{r dataNameMap{{this_data_name}}, fig.cap = paste0("Location of sites at the beginning of the cleaning process from ",use_value,", including the year of last visit. For definitions of regions, see (RSCA)[#rsca]")}

  m <- data_name_map(df = bio_df %>% filter(!is.na(lat))
                     , x = "long"
                     , y = "lat"
                     , polys = settings$aoi %>%
                       dplyr::mutate(polys = settings$name)
                     , aoi = settings$aoi
                     , polys_col = "polys"
                     , polys_name = settings$use_aoi
                     , polys_alpha = 0
                     , dots_args = list(title = paste0(use_value, " records")
                                        # , col = "year"
                                        , legend.format = list(big.mark = "")
                                        , palette = "viridis"
                                        , style = "fixed"
                                        , breaks = seq(1979, 2029, 10)
                                        , labels = paste0(seq(1980, 2020, 10)
                                                          , " to "
                                                          , seq(1989, 2029, 10)
                                                          ) %>% stringr::str_replace_all("2029"
                                                                                         , format(Sys.Date(),
                                                                                                  "%Y")
                                                                                         )
                                        )
                     )

  m

```

<br>

