```{r {{this_data_name}}}

  use_value <- "{{this_data_name}}" # use in file names; see loop at end of 0020_data.Rmd
  USE_VALUE <- toupper(use_value) # use in text

```

```{r dataSetup{{this_data_name}}}

    #find cached data
  bio_file <- fs::path(cache_dir, paste0("bio_", use_value, ".rds"))

  if(file.exists(bio_file)){
    
    load(bio_file)
    
  } else {
  
    #or collect source data from bio_all 
    bio_df <- arrow::open_dataset(bio_all_file) %>%
         dplyr::filter(year >= settings$min_year) %>%
         dplyr::filter(data_name == use_value) %>%
         dplyr::filter(original_name != "") %>% 
         dplyr::filter(!is.na(lat), !is.na(long)) %>% 
         dplyr::collect()
    
    save(bio_df, file = bio_file)
  
  }
      
  #plot/summaries options
  include_time_plot <- bio_df %>%
    dplyr::mutate(year = lubridate::year(date)
                  , month = lubridate::month(date)
    ) %>%
    dplyr::pull(year) %>%
    n_distinct() > 1
  
  include_summarise_cols <- all(c("original_name", "survey", "rel_metres_adj", "month", "year") %in% names(bio_df))
    

 
```

(ref:timePlotReference{{this_data_name}}) Figure \@ref(fig:timePlot{{this_data_name}})

(ref:dataNameMapReference{{this_data_name}}) Figure \@ref(fig:dataNameMap{{this_data_name}})

```{r dataNameSummary{{this_data_name}}, child = here::here("report", "child", "sources", paste0(use_value, ".Rmd"))}
```

#### Data


`r if(include_summarise_cols) {summarise_cols(bio_df, c("original_name", "survey", "kingdom", "rel_metres", "month", "year"), qs = c(0.05, 0.25, 0.5, 0.75, 0.95))$text}`

`r if(include_time_plot) {paste0("(ref:timePlotReference{{this_data_name}}) shows the distribution of records through time for ", USE_VALUE, ". ")} else paste(USE_VALUE, " did not have temporal data associated. ")`
`r if(include_dataNameMap) {paste0(" (ref:dataNameMapReference{{this_data_name}}) shows the spatial location of records for ", USE_VALUE, ", within South Australia.")}`

<br>

```{r timePlot{{this_data_name}}, fig.cap = paste("Records from", USE_VALUE, "through time"), eval = include_time_plot}

  data_name_time_plot(bio_df) +
    ggplot2::labs(subtitle = USE_VALUE)

```

`r if(include_time_plot) {"<br>"}`

```{r dataNameMap{{this_data_name}}, fig.cap = paste("Spatial location of records from", USE_VALUE), eval = include_dataNameMap}

tmap_mode("plot")

  m <- data_name_map(df = bio_df 
                     , x = "long"
                     , y = "lat"
                     , polys = settings$aoi %>%
                       dplyr::mutate(polys = settings$name)
                     , aoi = sa_nobuf_proj
                     , polys_col = "polys"
                     , polys_name = settings$use_aoi
                     , polys_alpha = 0
                     , dots_args = list(title = paste(USE_VALUE, " records")
                                        , col = "year"
                                        , legend.format = list(big.mark = "")
                                        , palette = "viridis"
                                        , style = "fixed"
                                        , breaks = seq(1969, 2029, 10)
                                        , labels = paste0(seq(1970, 2020, 10)
                                                          , " to "
                                                          , seq(1979, 2029, 10)
                                                          ) %>% stringr::str_replace_all("2029"
                                                                                         , format(Sys.Date(),
                                                                                                  "%Y")
                                                                                         )
                                        )
                     , layout_args = list(legend.outside = TRUE)
                     ) +
    tm_shape(sa_nobuf_proj) +
    tm_borders(col = "black")

 m
 
 tmap::tmap_mode(use_tmap_mode)
 
```

<br>
